---
title: "CRP Cleaning"
author: "César Higgins"
date: '2023-01-17'
output:  
    html_document:
    code_folding: hide
    highlight: pygments
    keep_md: no
    number_sections: yes
    smart: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---



```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  

summarize = dplyr::summarize

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```


# Dry Blood Sample hsCRP Data Cleaning 

## Importing Erin's Data Repository

First, import the data repository from Erin's database 

```{r read data, results="hide"}

rm(list = ls())
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
data_path  <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/"

paste0(source_dir, "hrs_democoggen_wide20200910.Rda")
load(paste0(source_dir, "hrs_democoggen_wide20200910.Rda")) 
dem <- hrs0910



```

Second, merge data repository with dry blood spot baseline samples 2006 and 2008 and extract all baseline biomarkers for each wave. The following biomarkers are in each data set: hsCRP, CysC, TC, HDL, and HbA1C. 


## Importing Dry Blood Spot (DBS)

```{r read data from DBS, results="hide"}

load(paste0(source_dir, "biomk06.RData"))
load(paste0(source_dir, "biomk08.RData"))
biomk06 <- biomk06
biomk08 <- biomk08

myvars_bio_2006 <- c("HHID", "PN", "KA1C_ADJ", "KHDL_ADJ", "KTC_ADJ", "KCYSC_ADJ", "KCRP_ADJ") 
myvars_bio_2008 <- c("HHID", "PN", "LA1C_ADJ", "LHDL_ADJ", "LTC_ADJ", "LCYSC_ADJ", "LCRP_ADJ") 

biomk06_f <- dplyr::select(biomk06,myvars_bio_2006)  
biomk08_f <- dplyr::select(biomk08,myvars_bio_2008)

# Renaming Variables for each data set biomarker data 
#2006 
biomk06_f <- dplyr::rename(biomk06_f, a1c_2006=KA1C_ADJ , hdl_2006 = KHDL_ADJ,  tc_2006 = KTC_ADJ, cysc_2006 = KCYSC_ADJ, crp_2006 = KCRP_ADJ)

# 2008 
biomk08_f <- dplyr::rename(biomk08_f, a1c_2008 = LA1C_ADJ, hdl_2008 = LHDL_ADJ, tc_2008 = LTC_ADJ,cysc_2008 = LCYSC_ADJ, crp_2008 = LCRP_ADJ)

```


# Mergin Biomarkers Data Set with Cognition Data Set 

The final data set has all 10 biomarkers, 5 from the year 2006 and 5 from the year 2008 

```{r}

merged_data <- left_join(dem, biomk06_f, by = c("HHID" = "HHID", "PN" = "PN")) 
merged_data_f <- left_join(merged_data, biomk08_f, by = c("HHID" = "HHID", "PN" = "PN"))

# Checking the dimensions of the data set to see that everything is fine 

dem %>% 
dim()

biomk06_f %>% 
dim()

biomk08_f %>% 
dim()

merged_data %>% 
dim()

merged_data_f %>% 
dim()
```


# Reducing the Number of Variables in my Data set

```{r}

myvars <- c("HHID", "PN", "a1c_2006", "hdl_2006", "tc_2006", "cysc_2006", "crp_2006", "a1c_2008", "hdl_2008", "tc_2008", "cysc_2008", "crp_2008", "DEGREE", "GENDER", "SCHLYRS", "HISPANIC", "RACE", "AD95", "AD96", "AD98", "AD00", "AD02", "AD04", "AD06", "AD08", "AD10", "AD12", "AD14", "expert_dem_98", "expert_dem_00", "expert_dem_02", "expert_dem_04", "expert_dem_06", "expert_dem_08", "expert_dem_10", "expert_dem_12", "expert_dem_14", "AAGE", "BAGE", "CAGE", "DAGE", "EAGE", "FAGE", "GAGE", "HAGE", "JAGE", "KAGE", "LAGE", "MAGE", "NAGE", "OAGE", "PAGE","totscore95", "totscore96", "totscore98", "totscore00", "totscore02", "totscore04", "totscore06", "totscore08", "totscore10", "totscore12", "totscore14", "R3HIBPE", "R4HIBPE", "R5HIBPE", "R6HIBPE", "R7HIBPE", "R8HIBPE", "R9HIBPE", "R10HIBPE", "R11HIBPE", "R12HIBPE", "smoke96", "smoke98", "smoke00", "smoke02", "smoke04", "smoke06", "smoke08", "smoke10", "smoke12", "smoke14","R3DIABE", "R4DIABE", "R5DIABE", "R6DIABE", "R7DIABE", "R8DIABE", "R9DIABE", "R10DIABE", "R11DIABE", "R12DIABE", "R3BMI", "R4BMI", "R5BMI", "R6BMI", "R7BMI", "R8BMI", "R9BMI", "R10BMI", "R11BMI", "R12BMI", "APOE2010_bin", "EA_APOE4", "AA_APOE4", "BIRTHYR", "STROKE95",  "STROKE96", "STROKE98", "STROKE00", "STROKE02", "STROKE04", "STROKE06", "STROKE08", "STROKE10", "STROKE12", "STROKE14", "DA009", "EA009", "FA009", "GA009", "HA009", "JA009", "KA009", "LA009", "MA009", "NA009", "OA009", "R8DRINKN", "R9DRINKN")

dem_final <- dplyr::select(merged_data_f,myvars)

dem_final %>% 
  dim()

```


# Creating the Outcome Variable

```{r}
# Data cleaning 

# Assigning Waves & Creating Outcome Variables 
dem_final_f <- dem_final %>% 
  mutate(wave_2006 = case_when(
    (!is.na(a1c_2006) | !is.na(hdl_2006) | !is.na(tc_2006) | !is.na(cysc_2006) | !is.na(crp_2006)) ~ 1, 
    TRUE ~ 0
  )) %>% 
  mutate(wave_2008 = case_when(
    (!is.na(a1c_2008) | !is.na(hdl_2008) | !is.na(tc_2008) | !is.na(cysc_2008) | !is.na(crp_2008)) ~ 1, 
    TRUE ~ 0
  )) %>% 
  mutate(wave = case_when(
    wave_2006==1 ~ 2006,
    wave_2008==1 ~ 2008,
  )) %>% 
  mutate(AD_06 = case_when(
    AD06==3 ~ 0, 
    AD06==2 ~ 1,
    AD06==1 ~ 2
  )) %>% 
  mutate(AD_08 = case_when(
    AD08==3 ~ 0,
    AD08==2 ~ 1, 
    AD08==1 ~ 2
  )) %>% 
  mutate(demdx= case_when(
    wave==2006 ~ AD_06,
    wave==2008 ~ AD_08)) 

```


# Covariates for Analysis 

Covariates include: 1) socio-demographics: age, gender, education, race
                    2) behavioral: BMI, alcohol, smoking 
                    3) mediators: tc, hdl, a1c, cysc, crp 
                    4) genetic marker: any apoe-e4 
      


```{r}

dem_final_ff <-  dem_final_f %>% 
  mutate(age = wave-BIRTHYR) %>% 
  mutate(education = SCHLYRS) %>% 
  mutate(stroke = case_when(
    wave==2006 ~ STROKE06,
    wave==2008 ~ STROKE08
  )) %>% 
  mutate(hbp = case_when(
    wave==2006 ~ R8HIBPE,
    wave==2008 ~ R9HIBPE 
  )) %>% 
  mutate(t2db = case_when(
    wave==2006 ~ R8DIABE,
    wave==2008 ~ R9DIABE
  )) %>% 
  mutate(bmi = case_when(
    wave==2006 ~ R8BMI,
    wave==2008 ~ R9BMI
  )) %>% 
  mutate(smoking = case_when(
    wave==2006 ~ smoke06,
    wave==2008 ~ smoke08
  )) %>% 
  mutate(alcohol = case_when(
    wave==2006 ~ R8DRINKN,
    wave==2008 ~ R9DRINKN
  )) %>% 
  mutate(proxy = case_when(
    wave==2006 ~ KA009,
    wave==2008 ~ LA009,
  )) %>% 
  mutate(proxy_f= case_when(
    proxy==1 ~ 1,
    proxy==2 ~ 2, 
    proxy==3 ~ 2
  )) %>% 
  mutate(totalscore= case_when(
    wave==2006 ~ totscore06,
    wave==2008 ~ totscore08
  )) %>% 
  mutate(race_f = case_when(
    RACE==0 ~ "NA",
    RACE==1 ~ "White",
    RACE==2 ~ "Black",
    RACE==7 ~ "Other")) %>% 
  mutate(hispanic_f = case_when(
      HISPANIC==1 | HISPANIC==2 | HISPANIC==3 ~ "Hispanic",
      HISPANIC==5 ~ "Non-Hispanic",
      HISPANIC==0 ~ "NA")) %>% 
  mutate(baseline_crp = case_when(
    wave==2006 ~ crp_2006,
    wave==2008 ~ crp_2008
  )) %>% 
  mutate(baseline_hdl= case_when(
    wave==2006 ~ hdl_2006,
    wave==2008 ~ hdl_2008
  )) %>% 
  mutate(baseline_cysc = case_when(
    wave==2006 ~ cysc_2006,
    wave==2008 ~ cysc_2008
  )) %>% 
  mutate(baseline_tc = case_when(
    wave==2006 ~ tc_2006,
    wave==2008 ~ tc_2008
  )) %>% 
  mutate(baseline_a1c = case_when(
    wave==2006 ~ a1c_2006,
    wave==2008 ~ a1c_2008
  ))

 dem_final_fff <- dem_final_ff %>% 
  mutate(race_f = na_if(race_f, "NA")) %>% 
  mutate(hispanic_f = na_if(hispanic_f, "NA")) %>% 
  mutate(race_ff = case_when(
    race_f=="White" & hispanic_f=="Non-Hispanic" ~ "White-NH",
    race_f=="Black" & hispanic_f=="Non-Hispanic" ~ "Black-NH",
    race_f=="Other" & hispanic_f=="Non-Hispanic" ~ "Other-NH",
    hispanic_f=="Hispanic" ~ "Hispanic")) %>% 
  mutate(gender_f = case_when(
    GENDER==1 ~ "Male",
    GENDER==2 ~ "Female")) %>% 
  mutate(edu = case_when(
    DEGREE==0 | DEGREE==1 | DEGREE==2 ~ "HS or <",
    DEGREE==3 | DEGREE==4 | DEGREE==9 ~ "College/Some",
    DEGREE==5 | DEGREE==6 ~ "> College")) 
 
myvars_2 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "stroke", "hbp", "t2db", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx", "APOE2010_bin") 

dem_data_ffff <- dplyr::select(dem_final_fff,myvars_2)  


```



# Sample Size 

```{r}

complete_data <- dem_data_ffff

complete_crp <- complete_data %>% 
  filter(baseline_crp!="NA")  # N = 12,394, we lost 670 from the 13,064
 
complete_hdl <- complete_crp %>% 
  filter(baseline_hdl!="NA") # N = 10,458 

complete_tc <- complete_hdl %>% 
   filter(baseline_tc!="NA") # N= 10,455

complete_cysc <- complete_tc %>% 
   filter(baseline_cysc!="NA") # N= 10,403

complete_a1c <- complete_cysc %>% 
  filter(baseline_a1c!="NA") # N=10,380

complete_data_biomarkers <- complete_data %>% 
filter(baseline_crp!="NA" & baseline_hdl!="NA" & baseline_tc!="NA" & baseline_cysc!="NA" & baseline_a1c!="NA") # 10,380 same as complete_a1c, therefore we lost 2,694 without complete biomarker data 


complete_demographics <- complete_data_biomarkers %>% 
  filter(race_ff!="NA" & gender_f!="NA" & education!="NA" & age!="NA")  # N = 10,362 , we passed from 10380 to 10362, we lost 18 people 

complete_health <- complete_demographics %>% 
  filter(alcohol!="NA", hbp!="NA", bmi!="NA", smoking!="NA") # Remove from here DB and Stroke, and leave only those with behavioral conditions and HBP N = 10,154, we lost 208

complete_outcomes <- complete_health %>%  
  filter(demdx!="NA" & totalscore!="NA") # N = 10,154

complete_genetics <- complete_outcomes %>% 
  filter(APOE2010_bin!="NA") # 9,305, we lost 849 without complete outcome data 


# Using na.omit and not including the mutually exclusive outcomes variable only demdx and totalscore 

myvars_3 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "hbp", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx") 

complete_data_filter <- dplyr::select(complete_data,myvars_3)  

data_nm <- complete_data_filter%>% na.omit() # I get the same number of observation with full sample 10,154

# Sample Including Genetics APOE-information 

myvars_4 <- c("HHID", "PN","race_ff", "gender_f", "edu", "education", "wave", "age", "alcohol", "hbp", "bmi", "smoking", "alcohol", "proxy_f", "totalscore", "baseline_crp", "baseline_hdl", "baseline_tc", "baseline_cysc", "baseline_a1c", "demdx", "APOE2010_bin")

complete_data_filter_apoe <- dplyr::select(complete_data,myvars_4)  

data_nm_apoe <- complete_data_filter_apoe%>% na.omit() # I get full sample 9,305 , which is like losing 3769 observation from the DBS sample size


#save(data_nm, file = paste0(data_path, "full_data_hsCRP.RData"))
#save(data_nm_apoe, file = paste0(data_path, "full_data_hsCRP_apoe.RData"))



```


# Adding Chronic Conditions for hsCRP data sets 

```{r}

rm(list = ls())
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
data_path  <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/"


load(paste0(source_dir, "full_data_hsCRP.RData")) 
load(paste0(source_dir, "full_data_hsCRP_apoe.RData")) 
load(paste0(source_dir, "randhrs1992_2018v1.rda")) 
rand_data <- randhrs1992_2018v1

myvars_rand <- c("HHID", "PN", "R8HEARTE", "R8STROKE", "R8PSYCHE", "R8ARTHRE", "R8LUNGE", "R8CANCRE", "R8DIABE", "R8CONDE", "R9HEARTE", "R9STROKE", "R9PSYCHE", "R9ARTHRE", "R9LUNGE", "R9CANCRE", "R9DIABE", "R9CONDE")

rand_data <- dplyr::select(rand_data,myvars_rand) 


# merged hsCRP DBS with rand variable for Chronic conditions, now create variable names for this chronic conditions       

hsCRP_data_dbs <- left_join(data_nm, rand_data, by = c("HHID" = "HHID", "PN" = "PN")) 

hsCRP_data_dbs <- hsCRP_data_dbs %>%    
                  mutate(stroke = case_when(
                  wave==2006 ~ R8STROKE,
                  wave==2008 ~ R9STROKE
                  )) %>% 
                  mutate(heart = case_when(
                  wave==2006 ~ R8HEARTE,
                  wave==2008 ~ R9HEARTE
                  )) %>% 
                  mutate(artr = case_when(
                  wave==2006 ~ R8ARTHRE,
                  wave==2008 ~ R9ARTHRE
                  )) %>% 
                  mutate(psych = case_when(
                  wave==2006 ~ R8PSYCHE,
                  wave==2008 ~ R9PSYCHE
                  )) %>%
                  mutate(lung = case_when(
                  wave==2006 ~ R8LUNGE,
                  wave==2008 ~ R9LUNGE
                  )) %>%
                  mutate(cancer = case_when(
                  wave==2006 ~ R8CANCRE,
                  wave==2008 ~ R9CANCRE
                  )) %>%
                  mutate(t2db = case_when(
                  wave==2006 ~ R8DIABE,
                  wave==2008 ~ R9DIABE
                  )) 
    
hsCRP_data_dbs_nm <- hsCRP_data_dbs %>% na.omit()

hsCRP_data_dbs_apoe <- left_join(data_nm_apoe, rand_data, by = c("HHID" = "HHID", "PN" = "PN")) 

hsCRP_data_dbs_apoe <- hsCRP_data_dbs_apoe %>%    
                  mutate(stroke = case_when(
                  wave==2006 ~ R8STROKE,
                  wave==2008 ~ R9STROKE
                  )) %>% 
                  mutate(heart = case_when(
                  wave==2006 ~ R8HEARTE,
                  wave==2008 ~ R9HEARTE
                  )) %>% 
                  mutate(artr = case_when(
                  wave==2006 ~ R8ARTHRE,
                  wave==2008 ~ R9ARTHRE
                  )) %>% 
                  mutate(psych = case_when(
                  wave==2006 ~ R8PSYCHE,
                  wave==2008 ~ R9PSYCHE
                  )) %>%
                  mutate(lung = case_when(
                  wave==2006 ~ R8LUNGE,
                  wave==2008 ~ R9LUNGE
                  )) %>%
                  mutate(cancer = case_when(
                  wave==2006 ~ R8CANCRE,
                  wave==2008 ~ R9CANCRE
                  )) %>%
                  mutate(t2db = case_when(
                  wave==2006 ~ R8DIABE,
                  wave==2008 ~ R9DIABE
                  ))

hsCRP_data_dbs_apoe_nm <- hsCRP_data_dbs_apoe %>% na.omit()


#save(hsCRP_data_dbs_nm, file = paste0(data_path, "hsCRP_DBS_RAND_baseline.RData"))
#save(hsCRP_data_dbs_apoe_nm, file = paste0(data_path, "hsCRP_DBS_RAND_baseline_apoe.RData"))



```


# Adding follow-up cognitive data to data set 

```{r}


rm(list = ls())
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
data_path  <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/"



load(paste0(source_dir, "hsCRP_DBS_RAND_baseline_apoe.RData"))
load(paste0(source_dir, "hrs_democoggen_wide20200910.Rda")) 
dementia <- hrs0910

# obtain follow-up cognitive variables
myvars <- c("HHID", "PN", "AD08", "AD10", "AD12", "AD14", "expert_dem_08", "expert_dem_10", "expert_dem_12", "expert_dem_14", "totscore08", "totscore10", "totscore12", "totscore14")

cognitive_followup_data <- dplyr::select(dementia,myvars)


crp_data_incidence <- left_join(hsCRP_data_dbs_apoe_nm, cognitive_followup_data, by = c("HHID" = "HHID", "PN" = "PN")) 


crp_data_incidence <- crp_data_incidence %>% 
                      mutate(dementia_b = case_when(
                        demdx==2 ~ 1,
                        demdx==0 ~ 0
                      )) %>% 
                      mutate(cind_b = case_when(
                        demdx==1 ~ 1,
                        demdx==0 ~ 0
                      ))

# write data for analysis into STATA
#library(haven)
#write_dta(crp_data_incidence, paste0(source_dir, "crp_data_incidence.dta"))

```






