---
title: "CRP Tables"
author: "César Higgins"
date: '2023-01-19'
output:
  html_document: null
  highlight: pygments
  word_document: default
  keep_md: no
  number_sections: yes
  smart: yes
  theme: sandstone
  toc: yes
  toc_float: yes
  code_folding: hide
---


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  
library(haven)

summarize = dplyr::summarize

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```



# Import Data 

After Stata Analysis, imported file to create tables for descriptive statistics 

```{r}

# import data from STATA crp_data_analysis.dta for data analysis and creation of descriptive tables 
rm(list = ls())
source_dir_dta <- "/Users/cesarhiggins/Documents/CRP Dementia/STATA/"
dementia_inc_final <- read_dta(paste0(source_dir_dta, "crp_data_incidence_analysis.dta")) 

```


# Table 1 by Cognitive Status{.tabset}

Distribution of baseline sample characteristics by dementia status after 6-year of follow-up, United States Health and Retirement Study, 2006 and 2008


```{r}
# Table 1


table1_data <- dementia_inc_final %>%  
  dplyr::select("dementia", "race", "sex", "edu_cat", "education", "age", "alcohol", "smoke", "bmi", "chronic", "baseline_crp", "APOE2010_bin", "t_dementia", "wave")

table1_data <- table1_data %>%  
                mutate(dementia_f = case_when(
                  dementia==0 ~ "Normal", 
                  dementia==1 ~ "Dementia"
                )) %>% 
                mutate(apoe4 = case_when(
                APOE2010_bin==0 ~ "No copy",
                APOE2010_bin==1 ~ "At least 1 copy"
                )) %>% 
                mutate_at(vars(apoe4), as.factor) %>% 
                mutate(race_f = case_when(
                  race==0 ~ "White-NH", 
                  race==1 ~ "Black-NH", 
                  race==2 ~ "Hispanic"
                )) %>%  
                mutate(chronic_f = case_when(
                  chronic==0 ~ "No", 
                  chronic>=1 ~ "At least 1"
                )) %>% 
                mutate(edu_cat_f = case_when(
                  edu_cat==1 ~ "> College",
                  edu_cat==2 ~ "College/Some",
                  edu_cat==3 ~ "HS or <"
                )) %>% 
                mutate(sex_f = case_when(
                  sex==1 ~ "Female", 
                  sex==2 ~ "Male"
                )) %>% 
                mutate(smoke_f = case_when(
                  smoke==0 ~ "Never", 
                  smoke==1 ~ "Current/Former"
                )) %>% 
                mutate(time = t_dementia) %>% 
                mutate_at(vars(time), as.numeric) %>% 
                mutate_at(vars(chronic), as.numeric)

table1_data <- table1_data %>%  
  dplyr::select("dementia_f", "baseline_crp", "age", "race_f", "sex_f", "edu_cat_f", "alcohol", "smoke_f", "bmi", "chronic_f", "apoe4", "time", "wave", "chronic")

table1_data_nm <- table1_data %>% na.omit() 


table1_cognition <- table1_data_nm %>%
  tbl_summary(
    by = dementia_f, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(
                baseline_crp ~ "hsCRP", 
                age ~ "Age",
                race_f ~ "Race",
                sex_f ~ "Gender", 
                edu_cat_f ~ "Educational Category", 
                alcohol ~ "Alcohol Use", 
                smoke_f ~ "Smoking",
                bmi ~ "Body Mass Index", 
                chronic_f ~ "Chronic Conditions", 
                apoe4 ~ "APOE-e4", 
                time ~ "Time", 
                wave ~ "Wave")) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table1_cognition


# Save analytic sample size data 
# save(table1_data_nm, file=paste0(source_dir_dta, "Complete_hsCRP.RData")) 


```


# Supplemental Table 4{.tabset}

Supplemental Table 4: Distribution of baseline sample characteristics by percentiles of C-reactive protein, United States Health and Retirement Study, 2006 and 2008

```{r}


table1_data_nm %>% 
  summarize("Obs" = n(),
  "Mean" =mean(baseline_crp,na.rm=TRUE),
  "SD" = sd(baseline_crp,na.rm=TRUE),
  "Min" = min(baseline_crp, na.rm=TRUE),
  "Max" = max(baseline_crp, na.rm=TRUE),
  "25%" = quantile(baseline_crp, probs=0.25, na.rm=TRUE), "50%" = quantile(baseline_crp, probs=0.50, na.rm=TRUE), "75%" = quantile(baseline_crp, probs=0.75, na.rm=TRUE),
  "IQR" = IQR(baseline_crp, na.rm=TRUE)) %>% 
  na.omit() 


table2_data <- dementia_inc_final %>% 
  mutate(crp_cat = case_when(
    baseline_crp < 0.96 ~ 0, 
    baseline_crp >= 0.96 & baseline_crp < 1.97 ~ 1,
    baseline_crp >= 1.97 & baseline_crp < 4.57 ~ 2, 
    baseline_crp >= 4.57 ~ 3
  )) %>% 
  mutate(dementia_f = case_when(
                  dementia==0 ~ "Normal", 
                  dementia==1 ~ "Dementia"
                )) %>% 
                mutate(apoe4 = case_when(
                APOE2010_bin==0 ~ "No copy",
                APOE2010_bin==1 ~ "At least 1 copy"
                )) %>% 
                mutate_at(vars(apoe4), as.factor) %>% 
                mutate(race_f = case_when(
                  race==0 ~ "White-NH", 
                  race==1 ~ "Black-NH", 
                  race==2 ~ "Hispanic"
                )) %>%  
                mutate(chronic_f = case_when(
                  chronic==0 ~ "No", 
                  chronic>=1 ~ "At least 1"
                )) %>% 
                mutate(edu_cat_f = case_when(
                  edu_cat==1 ~ "> College",
                  edu_cat==2 ~ "College/Some",
                  edu_cat==3 ~ "HS or <"
                )) %>% 
                mutate(sex_f = case_when(
                  sex==1 ~ "Female", 
                  sex==2 ~ "Male"
                )) %>% 
                mutate(smoke_f = case_when(
                  smoke==0 ~ "Never", 
                  smoke==1 ~ "Current/Former"
                )) %>% 
                mutate(time = t_dementia) %>% 
                mutate_at(vars(time), as.numeric) %>% 
                mutate_at(vars(chronic), as.numeric)

table2_data <- table2_data %>%  
  dplyr::select("dementia_f", "crp_cat", "age", "race_f", "sex_f", "edu_cat_f", "alcohol", "smoke_f", "bmi", "chronic_f", "apoe4", "time", "wave")

table2_data_nm <- table2_data %>% na.omit()

table2_crp <- table2_data_nm %>%
  tbl_summary(
    by = crp_cat, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(
                dementia_f ~ "Dementia",
                age ~ "Age",
                race_f ~ "Race",
                sex_f ~ "Gender", 
                edu_cat_f ~ "Educational Category", 
                alcohol ~ "Alcohol Use", 
                smoke_f ~ "Smoking",
                bmi ~ "Body Mass Index", 
                chronic_f ~ "Chronic Conditions", 
                apoe4 ~ "APOE-e4", 
                time ~ "Time", 
                wave ~ "Wave")) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table2_crp


```


# Supplemental Table 2{.tabset}

Supplemental Table 2: Distribution of baseline sample characteristics by racialized social groups in the Health and Retirement Study, 2006 and 2008

```{r}

table3_data <- dementia_inc_final %>% 
                mutate(dementia_f = case_when(
                  dementia==0 ~ "Normal", 
                  dementia==1 ~ "Dementia"
                )) %>% 
                mutate(apoe4 = case_when(
                APOE2010_bin==0 ~ "No copy",
                APOE2010_bin==1 ~ "At least 1 copy"
                )) %>% 
                mutate_at(vars(apoe4), as.factor) %>% 
                mutate(race_f = case_when(
                  race==0 ~ "White-NH", 
                  race==1 ~ "Black-NH", 
                  race==2 ~ "Hispanic"
                )) %>%  
                mutate(chronic_f = case_when(
                  chronic==0 ~ "No", 
                  chronic>=1 ~ "At least 1"
                )) %>% 
                mutate(edu_cat_f = case_when(
                  edu_cat==1 ~ "> College",
                  edu_cat==2 ~ "College/Some",
                  edu_cat==3 ~ "HS or <"
                )) %>% 
                mutate(sex_f = case_when(
                  sex==1 ~ "Female", 
                  sex==2 ~ "Male"
                )) %>% 
                mutate(smoke_f = case_when(
                  smoke==0 ~ "Never", 
                  smoke==1 ~ "Current/Former"
                )) %>% 
                mutate(time = t_dementia) %>% 
                mutate_at(vars(time), as.numeric) %>% 
                mutate_at(vars(chronic), as.numeric)

table3_data <- table3_data %>%  
  dplyr::select("race_f", "dementia_f", "baseline_crp", "age", "sex_f", "edu_cat_f", "alcohol", "smoke_f", "bmi", "chronic_f", "apoe4", "time", "wave")

table3_data_nm <- table3_data %>% na.omit()


table3_race <- table3_data_nm %>%
  tbl_summary(
    by = race_f, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(
                dementia_f ~ "Dementia",
                baseline_crp ~ "hsCRP", 
                age ~ "Age",
                sex_f ~ "Gender", 
                edu_cat_f ~ "Educational Category", 
                alcohol ~ "Alcohol Use", 
                smoke_f ~ "Smoking",
                bmi ~ "Body Mass Index", 
                chronic_f ~ "Chronic Conditions", 
                apoe4 ~ "APOE-e4", 
                time ~ "Time", 
                wave ~ "Wave")) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table3_race



```



# Supplemental Table 3 & Supplemental Figure 2{.tabset}

Supplemental Table 3 & Supplemental Figure 2: Baseline levels of C-reactive protein (CRP) by racialized social groups and sex in the United States Health and Retirement Study, waves 2006 & 2008


```{r}

crp_race <- table1_data_nm %>% 
  group_by(race_f, sex_f) %>% 
  summarize("Obs" = n(),
  "Mean" =mean(baseline_crp,na.rm=TRUE),
  "SD" = sd(baseline_crp,na.rm=TRUE),
  "Min" = min(baseline_crp, na.rm=TRUE),
  "Max" = max(baseline_crp, na.rm=TRUE),
  "25%" = quantile(baseline_crp, probs=0.25, na.rm=TRUE), "50%" = quantile(baseline_crp, probs=0.50, na.rm=TRUE), "75%" = quantile(baseline_crp, probs=0.75, na.rm=TRUE),
  "IQR" = IQR(baseline_crp, na.rm=TRUE)) %>% 
  na.omit() 

crp_race <- crp_race %>% 
            mutate(race_sex = case_when(
              race_f=='Black-NH' & sex_f=='Female' ~ "Black-NH Female",
              race_f=='Black-NH' & sex_f=='Male' ~ "Black-NH Male",
              race_f=='Hispanic' & sex_f=='Female' ~ "Hispanic Female",
              race_f=='Hispanic' & sex_f=='Male' ~ "Hispanic Male",
              race_f=='White-NH' & sex_f=='Female' ~ "White-NH Female",
              race_f=='White-NH' & sex_f=='Male' ~ "White-NH Male"
            )) %>% 
            mutate(crp=Mean) %>% 
            mutate(LL=Mean-SD) %>% 
            mutate(UL=Mean+SD)


crp_race 





table_base <- ggplot(crp_race, aes(y = race_sex, x = crp)) +
  geom_point(shape = 18, size = 5) +  
  geom_errorbarh(aes(xmin = LL, xmax = UL), height = 0.25) +
  geom_vline(xintercept = 4.18, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
  xlab("Mean C-reactive protein levels (\u03bcg/mL)") + 
  ylab(" ") + 
   theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 25, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"))

table_base


tab1 <- table_base + 
  labs(title = "space") +
  geom_text(aes(y = race_sex, x = -10, label = sprintf("%0.1f", round(crp, digits = 2))), size = 4) + ## decimal places
  ggtitle("Mean")


```



# Supplemental Table 1{.tabset}

Supplemental Table 1: Distribution of baseline characteristics between included and excluded participants in the United States Retirement Study 


```{r}

# First import analytic sample size and the one that we would like to compare with missing covatiates 

rm(list = ls())
source_dir <- "/Users/cesarhiggins/Documents/Dementia project /NIH grant/Data/" 
load(paste0(source_dir, "hsCRP_DBS_RAND_baseline_apoe.RData")) # file with all info on baseline covariates 

source_dir_dta <- "/Users/cesarhiggins/Documents/CRP Dementia/STATA/"
dementia_inc_final <- read_dta(paste0(source_dir_dta, "crp_data_incidence_analysis.dta")) # file with analytic sample size

baseline_data <- hsCRP_data_dbs_apoe_nm

# Create a dummy variable in the  in the analytic sample, because everyone has values for all covariates missing==0

dementia_inc_final <- dementia_inc_final %>% 
                          mutate(missing=0)

myvars_not_missing <- c("HHID", "PN", "missing", "dementia") 

dementia_inc_final <- dplyr::select(dementia_inc_final, myvars_not_missing) 

data_mising_table <- left_join(baseline_data, dementia_inc_final, by = c("HHID" = "HHID", "PN" = "PN"))

data_mising_table <- data_mising_table %>% 
  mutate(missing_f = ifelse(is.na(missing), 1, 0))



```


## Table of Missing Data

```{r}

data_mising_table <- data_mising_table %>% 
  mutate(dementia_f = case_when(
    dementia==0  ~ "normal",
    dementia==1 ~ "dementia"
  )) %>%  
  mutate(smoking_f = case_when(
    smoking==0 ~ "never",
    smoking==1 | smoking==2 ~ "current/former"
  )) %>% 
  mutate(chronic = case_when(
    wave==2006 ~ R8CONDE,
    wave==2008 ~ R9CONDE
  )) %>% 
  mutate(chronic_f = case_when(
    chronic>=1 ~ "At least 1",
    chronic==0 ~ "No Chronic"
  )) %>% 
  mutate(apoe = case_when(
    APOE2010_bin==1 ~ "At least 1",
    APOE2010_bin==0 ~ "No copy"
  )) 
  
# NOTE: this table does not need to include dementia status because dementia was assessed prospectively and of course nobody in the excluded sample would have a diagnosis of incident dementia

table1_excluded <- data_mising_table %>%  
  dplyr::select("dementia_f", "race_ff", "gender_f", "edu","age", "alcohol", "smoking_f", "bmi", "baseline_crp", "wave" , "missing_f", "apoe", "chronic_f")

table_missing <- table1_excluded %>%
  tbl_summary(
    by = missing_f, missing = "no",
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}  ({p}%)"),
    digits = all_continuous() ~ 2,
     label = list(
                  dementia_f ~ "Cognitive Status",
                  baseline_crp ~ "Baseline hsCRP",
                  age ~ "Age",
                  race_ff ~ "Race",
                  gender_f ~ "Gender", 
                  edu ~ "Education Category",
                  alcohol ~ "Alcohol",
                  smoking_f ~ "Smoking",
                  bmi ~ "Body Mass Index",
                  chronic_f ~ "Chronic Conditions",
                  apoe ~ "APOE-e4",
                  wave ~ "Wave"
  )) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "aov"),
        pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%  
  bold_p() %>%  
  bold_labels() 

table_missing %>% as_flex_table

```




# Figure 2 & Supplemental Table 1{.tabset}

Supplemental Figure 4: Density plot of natural logarithmic transformation of C-reactive protein (CRP) in a selected sample of United States adults in the Health and Retirement Study.             A. Distribution of the natural logarithmic transformation of C-reactive protein (CRP) by APOE-4 allele carrier status in our selected sample from the Health and Retirement Study (HRS). Dotted line denotes the cut off point for elevated levels of CRP at the 75th percentile (> 4.57μg/mL).
B. Distribution of CRP by racialized groups and APOE-4 allele carrier status. White-NH: non-Hispanic White, Black-NH: non-Hispanic Black.


```{r}

library(tidyverse)     ## data wrangling + ggplot2
library(colorspace)    ## adjust colors
library(rcartocolor)   ## Carto palettes
library(ggforce)       ## sina plots
library(ggdist)        ## halfeye plots
library(ggridges)      ## ridgeline plots
library(ggbeeswarm)    ## beeswarm plots
library(gghalves)      ## off-set jitter
library(systemfonts)   ## custom fonts
library(tidyquant)
library(ggpubr)

rm(list = ls())
source_dir_dta <- "/Users/cesarhiggins/Documents/CRP Dementia/STATA/"
dementia_inc_final <- read_dta(paste0(source_dir_dta, "crp_data_incidence_analysis.dta")) 

table1_data <- dementia_inc_final %>%  
  dplyr::select("dementia", "race", "sex", "edu_cat", "education", "age", "alcohol", "smoke", "bmi", "chronic", "baseline_crp", "APOE2010_bin", "t_dementia", "wave")


table1_data <- table1_data %>% 
                  mutate(race_f = case_when(
                  race==0 ~ "White-NH", 
                  race==1 ~ "Black-NH", 
                  race==2 ~ "Hispanic"
                )) %>% 
                mutate(apoe4 = case_when(
                APOE2010_bin==0 ~ "No copy",
                APOE2010_bin==1 ~ "At least 1 copy"
                ))




# Figure 2

figure_1 <- table1_data %>%
  ggplot(aes(y = race_f, x = log(baseline_crp), fill=race_f)) +
  stat_halfeye(justification = -.05) +
  geom_vline(xintercept = log(4.57), color = "gray85", linetype = "dashed") +
  scale_fill_tq() +
  theme_classic2() +
      labs(
    x = "Natural log transformation of C-reactive protein",
    y = "Racilized Groups",
    fill = "Racilized Groups"
  )

figure_1




# Supplemental Figure 3
## APOE-e4 Figures 

figure_2a <- table1_data %>%
  ggplot(aes(y = apoe4, x = log(baseline_crp), fill=apoe4)) +
  stat_halfeye(justification = -.05) +
  geom_vline(xintercept = log(4.57), color = "gray85", linetype = "dashed") +
  scale_fill_tq() +
  theme_classic2() +
      labs(
    x = "Natural log transformation of C-reactive protein",
    y = "APOE-\u03B5 4",
    fill = "APOE-\u03B5 4"
  )

figure_2a


figure_2b <- table1_data %>%
  ggplot(aes(y = factor(race_f), x = log(baseline_crp))) +
  stat_halfeye(
    aes(fill = apoe4),
   justification = -.05, position = "dodgejust"
  ) +
  geom_vline(xintercept = log(4.57), color = "gray85", linetype = "dashed") +
  labs(
      x = "Natural log transformation of C-reactive protein",
    y = "Racialized Groups",
    fill = "APOE-\u03B5 4"
  ) +
  scale_fill_tq() +
  theme_classic2() 

figure_2b

figures_2 <- ggarrange(
  figure_2a, figure_2b, labels = c("A", "B"),
  common.legend = TRUE, legend = "right"
  )
figures_2 


```



