---
title: "Mediation CRP Analysis"
author: "César Higgins"
date: '2023-09-05'
output:
  html_document: null
  highlight: pygments
  word_document: default
  keep_md: no
  number_sections: yes
  smart: yes
  theme: sandstone
  toc: yes
  toc_float: yes
  code_folding: hide
---


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  
library(haven)

summarize = dplyr::summarize
select = dplyr::select

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```




# Mediation 

```{r}

library(haven)
# import data from stata for mediation analysis
rm(list = ls())
source_dir_dta <- "/Users/cesarhiggins/Documents/CRP_Dementia_8-9-23/STATA/"
dementia_incidence_6y <- read_dta(paste0(source_dir_dta, "crp_data_incidence_mediation_9-5-23.dta")) # file with analytic sample size


library(CMAverse)

```


```{r, results='hide'}

library(writexl)
source_dir_data <- "/Users/cesarhiggins/Documents/CRP_Dementia_8-9-23/Mediation Results/"

# function to extract decomposition components 
# Link to existing files for function use
source(paste0(source_dir_data, "extract_decompo.R"))
source(paste0(source_dir_data, "extract_evalue.R"))

```



# Black Models

```{r, results='hide'}


select = dplyr::select 

black_data_dementia <- dementia_incidence_6y %>% dplyr::select(dementia, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()

black_data_dementia_2 <- dementia_incidence_6y %>% select(dementia_2, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

# this last data set has N=5,666. Because there were 1,515 people with dementia/cind at baseline that were excluded from the data set, their time to dementia will be 0, and when we take the log, their value will go to NA. The cross-tab in stata of tab black dementia_3 if t_dementia_3!=0 will show that the N of this sample is 5,666. Just what we get in R

black_data_dementia_3 <- dementia_incidence_6y %>% select(dementia_3, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_3) %>% na.omit()


# Add Dataset for log-transformed and standardize C-reactive protein 

black_data_dementia_logcrp <- dementia_incidence_6y %>% select(dementia_2, black, baseline_crp, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()


# Regression Based Approach for dementia cases only 
set.seed(2021)
res_rb_b_1 <- cmest(data = black_data_dementia, model = "rb", outcome = "dementia", exposure = "black",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), EMint=TRUE,
                mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia)), yreg = glm(dementia ~ black + baseline_crp_75 + black*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=black_data_dementia),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach for dementia cases when cind is included in the normal category 
set.seed(2021)
res_rb_b_2 <- cmest(data = black_data_dementia_2, model = "rb", outcome = "dementia_2", exposure = "black",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), EMint=TRUE,
                mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_2)), yreg = glm(dementia_2 ~ black + baseline_crp_75 + black*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=black_data_dementia_2),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach for dementia cases when cind is included in the dementia category 
set.seed(2021)
res_rb_b_3 <- cmest(data = black_data_dementia_3, model = "rb", outcome = "dementia_3", exposure = "black",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_3"), EMint=TRUE,
                mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_3)), yreg = glm(dementia_3 ~ black + baseline_crp_75 + black*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=black_data_dementia_3),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)



# Regression Based Approach for dementia cases when cind is included in the normal category and crp is log transformed

black_data_dementia_logcrp <- black_data_dementia_logcrp %>% 
                              mutate(baseline_crp_log = log(baseline_crp)) %>% 
                              mutate(baseline_crp_z = (baseline_crp_log - mean(baseline_crp_log))/sd(baseline_crp_log))
black_data_dementia_logcrp <- black_data_dementia_logcrp %>% 
                              select(-c(baseline_crp, baseline_crp_log))


set.seed(2021)
res_rb_b_4 <- cmest(data = black_data_dementia_logcrp, model = "rb", outcome = "dementia_2", exposure = "black",
                mediator = c("baseline_crp_z"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), EMint=TRUE,
                mreg = list(glm(baseline_crp_z ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = gaussian(link = "identity"), data = black_data_dementia_logcrp)), yreg = glm(dementia_2 ~ black + baseline_crp_z + black*baseline_crp_z+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=black_data_dementia_logcrp),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# G estimation L approach for dementia cases only 

set.seed(2021)
res_gformula_b_l_1 <- cmest(data = black_data_dementia, model = "gformula", outcome = "dementia", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia)), yreg = glm(dementia ~ black + baseline_crp_75 + black*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=black_data_dementia),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach for dementia cases when cind is included in normal 


set.seed(2021)
res_gformula_b_l_2 <- cmest(data = black_data_dementia_2, model = "gformula", outcome = "dementia_2", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_2)), yreg = glm(dementia_2 ~ black + baseline_crp_75 + black*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=black_data_dementia_2),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_2)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach for dementia cases when cind is included in dementia 

set.seed(2021)
res_gformula_b_l_3 <- cmest(data = black_data_dementia_3, model = "gformula", outcome = "dementia_3", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_3"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_3)), yreg = glm(dementia_3 ~ black + baseline_crp_75 + black*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=black_data_dementia_3),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_3)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach for dementia cases when cind is included in normal and crp is log transformed


set.seed(2021)
res_gformula_b_l_4 <- cmest(data = black_data_dementia_logcrp, model = "gformula", outcome = "dementia_2", exposure = "black",
                      mediator = c("baseline_crp_z"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list(glm(baseline_crp_z ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = gaussian(link = "identity"), data = black_data_dementia_logcrp)), yreg = glm(dementia_2 ~ black + baseline_crp_z + black*baseline_crp_z + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=black_data_dementia_logcrp),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_2)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


```

## hsCRP Mediation Incident Dementia Blacks Models{.tabset}


### Regression Based Dementia_1 

```{r}

summary(res_rb_b_1)

result_black_crp_rb_1 = extract_decompo(res_rb_b_1)

ggcmest(res_rb_b_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_rb_1 <- cmsens(res_rb_b_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_1)

```

### Regression Based Dementia_2


```{r}

summary(res_rb_b_2)

result_black_crp_rb_2 = extract_decompo(res_rb_b_2)

ggcmest(res_rb_b_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_rb_2 <- cmsens(res_rb_b_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_2)




```


### Regression Based Dementia_3


```{r}

summary(res_rb_b_3)

result_black_crp_rb_3 = extract_decompo(res_rb_b_3)

ggcmest(res_rb_b_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_rb_3 <- cmsens(res_rb_b_3, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_3)

```


### Regression Based Dementia_4

```{r}

summary(res_rb_b_4)

result_black_crp_rb_4 = extract_decompo(res_rb_b_4)

ggcmest(res_rb_b_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_rb_4 <- cmsens(res_rb_b_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_4)

```


### Sensitivity G Method Dementia_1 

```{r}

summary(res_gformula_b_l_1)

result_black_crp_gformula_l_1 = extract_decompo(res_gformula_b_l_1)

ggcmest(res_gformula_b_l_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_gformula_b_l_1 <- cmsens(res_gformula_b_l_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_b_l_1)


```


### Sensitivity G Method Dementia_2

```{r}

summary(res_gformula_b_l_2)

result_black_crp_gformula_l_2 = extract_decompo(res_gformula_b_l_2)

ggcmest(res_gformula_b_l_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_gformula_b_l_2 <- cmsens(res_gformula_b_l_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_b_l_2)


```


### Sensitivity G Method Dementia_3

```{r}

summary(res_gformula_b_l_3)

result_black_crp_gformula_l_3 = extract_decompo(res_gformula_b_l_3)

ggcmest(res_gformula_b_l_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_gformula_b_l_3 <- cmsens(res_gformula_b_l_3, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_b_l_3)


```



### Sensitivity G Method Dementia_4

```{r}

summary(res_gformula_b_l_4)

result_black_crp_gformula_l_4 = extract_decompo(res_gformula_b_l_4)

ggcmest(res_gformula_b_l_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_gformula_b_l_4 <- cmsens(res_gformula_b_l_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_b_l_4)


```




# Hispanic Models


```{r, results='hide'}

hispanic_data_dementia <- dementia_incidence_6y %>% select(dementia, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()


hispanic_data_dementia_2 <- dementia_incidence_6y %>% select(dementia_2, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()


# this last data set has N=5,571. Because there were 1,515 people with dementia/cind at baseline that were excluded from the data set, their time to dementia will be 0, and when we take the log, their value will go to NA. The cross-tab in stata of tab hispanic dementia_3 if t_dementia_3!=0 will show that the N of this sample is 5,571. Just what we get in R

hispanic_data_dementia_3 <- dementia_incidence_6y %>% select(dementia_3, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_3) %>% na.omit()


hispanic_data_dementia_logcrp <- dementia_incidence_6y %>% select(dementia_2, hispanic, baseline_crp, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()


# Regression Based Approach dementia only 
set.seed(2021)
res_rb_h_1 <- cmest(data = hispanic_data_dementia, model = "rb", outcome = "dementia", exposure = "hispanic",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=hispanic_data_dementia),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach dementia including cind in normal category 
set.seed(2021)
res_rb_h_2 <- cmest(data = hispanic_data_dementia_2, model = "rb", outcome = "dementia_2", exposure = "hispanic",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_2"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_2),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach dementia including cind in normal category 
set.seed(2021)
res_rb_h_2 <- cmest(data = hispanic_data_dementia_2, model = "rb", outcome = "dementia_2", exposure = "hispanic",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_2"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_2),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach dementia including cind in dementia category 
set.seed(2021)
res_rb_h_3 <- cmest(data = hispanic_data_dementia_3, model = "rb", outcome = "dementia_3", exposure = "hispanic",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_3"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia_3 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=hispanic_data_dementia_3),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


hispanic_data_dementia_logcrp <- hispanic_data_dementia_logcrp %>% 
                              mutate(baseline_crp_log = log(baseline_crp)) %>% 
                              mutate(baseline_crp_z = (baseline_crp_log - mean(baseline_crp_log))/sd(baseline_crp_log))
hispanic_data_dementia_logcrp <- hispanic_data_dementia_logcrp %>% 
                              select(-c(baseline_crp, baseline_crp_log))

# Regression Based Approach dementia including cind in normal category and logcrp
set.seed(2021)
res_rb_h_4 <- cmest(data = hispanic_data_dementia_logcrp, model = "rb", outcome = "dementia_2", exposure = "hispanic",
                mediator = c("baseline_crp_z"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_2"), EMint = TRUE,
                mreg = list("linear"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_z + hispanic*baseline_crp_z+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_logcrp),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# G estimation L approach dementia cases only 

set.seed(2021)
res_gformula_h_l_1 <- cmest(data = hispanic_data_dementia, model = "gformula", outcome = "dementia", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=hispanic_data_dementia),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach dementia cases including cind in the normal category 

set.seed(2021)
res_gformula_h_l_2 <- cmest(data = hispanic_data_dementia_2, model = "gformula", outcome = "dementia_2", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_2),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia_2)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach dementia cases including cind in the normal category 

set.seed(2021)
res_gformula_h_l_3 <- cmest(data = hispanic_data_dementia_3, model = "gformula", outcome = "dementia_3", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_3"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_3 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=hispanic_data_dementia_3),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia_3)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


# G estimation L approach dementia cases including cind in the normal category 

set.seed(2021)
res_gformula_h_l_4 <- cmest(data = hispanic_data_dementia_logcrp, model = "gformula", outcome = "dementia_2", exposure = "hispanic",
                      mediator = c("baseline_crp_z"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("linear"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_z + hispanic*baseline_crp_z + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_logcrp),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia_logcrp)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


```


## hsCRP Mediation Incident Dementia Hispanic Models{.tabset}



### Regression Based Dementia # 1


```{r}

summary(res_rb_h_1)

result_hispanic_crp_rb_1 = extract_decompo(res_rb_h_1)

ggcmest(res_rb_h_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))



evalue_rb_h_1 <- cmsens(res_rb_h_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_h_1)

```

### Regression Based Dementia # 2


```{r}

summary(res_rb_h_2)

result_hispanic_crp_rb_2 = extract_decompo(res_rb_h_2)

ggcmest(res_rb_h_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))



evalue_rb_h_2 <- cmsens(res_rb_h_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_h_2)

```


### Regression Based Dementia # 3


```{r}

summary(res_rb_h_3)

result_hispanic_crp_rb_3 = extract_decompo(res_rb_h_3)

ggcmest(res_rb_h_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))



evalue_rb_h_3 <- cmsens(res_rb_h_3, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_h_3)

```


### Regression Based Dementia # 4


```{r}

summary(res_rb_h_4)

result_hispanic_crp_rb_4 = extract_decompo(res_rb_h_4)

ggcmest(res_rb_h_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))



evalue_rb_h_4 <- cmsens(res_rb_h_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_h_4)

```


### Sensitivity G Method Dementia_1 

```{r}

summary(res_gformula_h_l_1)

result_hispanic_crp_gformula_l_1 = extract_decompo(res_gformula_h_l_1)

ggcmest(res_gformula_h_l_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_h_l_1 <- cmsens(res_gformula_h_l_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_h_l_1)

```



### Sensitivity G Method Dementia_2 

```{r}

summary(res_gformula_h_l_2)

result_hispanic_crp_gformula_l_2 = extract_decompo(res_gformula_h_l_2)

ggcmest(res_gformula_h_l_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_h_l_2 <- cmsens(res_gformula_h_l_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_h_l_2)

```


### Sensitivity G Method Dementia_3 

```{r}

summary(res_gformula_h_l_3)

result_hispanic_crp_gformula_l_3 = extract_decompo(res_gformula_h_l_3)

ggcmest(res_gformula_h_l_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_h_l_3 <- cmsens(res_gformula_h_l_3, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_h_l_3)

```


### Sensitivity G Method Dementia_4 

```{r}

summary(res_gformula_h_l_4)

result_hispanic_crp_gformula_l_4 = extract_decompo(res_gformula_h_l_4)

ggcmest(res_gformula_h_l_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_h_l_4 <- cmsens(res_gformula_h_l_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_h_l_4)

```


# Minoritized Models 


```{r, results='hide'}

minority_data_dementia <- dementia_incidence_6y %>% select(dementia, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()

minority_data_dementia_2 <- dementia_incidence_6y %>% select(dementia_2, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

minority_data_dementia_3 <- dementia_incidence_6y %>% select(dementia_3, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_3) %>% na.omit()

minority_data_dementia_logcrp <- dementia_incidence_6y %>% select(dementia_2, minority, baseline_crp, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

# Regression Based Approach dementia 1
set.seed(2021)
res_rb_m_1 <- cmest(data = minority_data_dementia, model = "rb", outcome = "dementia", exposure = "minority",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia ~ minority + baseline_crp_75 + minority*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=minority_data_dementia),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach dementia 2
set.seed(2021)
res_rb_m_2 <- cmest(data = minority_data_dementia_2, model = "rb", outcome = "dementia_2", exposure = "minority",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_2"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia_2 ~ minority + baseline_crp_75 + minority*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=minority_data_dementia_2),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# Regression Based Approach dementia 3
set.seed(2021)
res_rb_m_3 <- cmest(data = minority_data_dementia_3, model = "rb", outcome = "dementia_3", exposure = "minority",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_3"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia_3 ~ minority + baseline_crp_75 + minority*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=minority_data_dementia_3),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


minority_data_dementia_logcrp <- minority_data_dementia_logcrp %>% 
                              mutate(baseline_crp_log = log(baseline_crp)) %>% 
                              mutate(baseline_crp_z = (baseline_crp_log - mean(baseline_crp_log))/sd(baseline_crp_log))
minority_data_dementia_logcrp <- minority_data_dementia_logcrp %>% 
                              select(-c(baseline_crp, baseline_crp_log))

# Regression Based Approach dementia 2 with log crp
set.seed(2021)
res_rb_m_4 <- cmest(data = minority_data_dementia_logcrp, model = "rb", outcome = "dementia_2", exposure = "minority",
                mediator = c("baseline_crp_z"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2" ,"chronic", "ln_time_2"), EMint = TRUE,
                mreg = list("linear"), yreg = glm(dementia_2 ~ minority + baseline_crp_z + minority*baseline_crp_z+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=minority_data_dementia_logcrp),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# G estimation L approach dementia_1

set.seed(2021)
res_gformula_m_l_1 <- cmest(data = minority_data_dementia, model = "gformula", outcome = "dementia", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=minority_data_dementia),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)

# G estimation L approach dementia_2

set.seed(2021)
res_gformula_m_l_2 <- cmest(data = minority_data_dementia_2, model = "gformula", outcome = "dementia_2", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_2 ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=minority_data_dementia_2),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia_2)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)

# G estimation L approach dementia_3

set.seed(2021)
res_gformula_m_l_3 <- cmest(data = minority_data_dementia_3, model = "gformula", outcome = "dementia_3", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_3"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_3 ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_3, family = poisson(link = "log"), data=minority_data_dementia_3),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia_3)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)

# G estimation L approach dementia_2 and log crp 

set.seed(2021)
res_gformula_m_l_4 <- cmest(data = minority_data_dementia_logcrp, model = "gformula", outcome = "dementia_2", exposure = "minority",
                      mediator = c("baseline_crp_z"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("linear"), yreg = glm(dementia_2 ~ minority + baseline_crp_z + minority*baseline_crp_z + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=minority_data_dementia_logcrp),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia_logcrp)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


```



## hsCRP Mediation Incident Dementia Minority Models{.tabset}

### Regression Based Dementia_1 

```{r}

summary(res_rb_m_1)

result_minority_crp_rb_1 = extract_decompo(res_rb_m_1)

ggcmest(res_rb_m_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_rb_m_1 <- cmsens(res_rb_m_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_m_1)

```


### Regression Based Dementia_2 

```{r}

summary(res_rb_m_2)

result_minority_crp_rb_2 = extract_decompo(res_rb_m_2)

ggcmest(res_rb_m_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_rb_m_2 <- cmsens(res_rb_m_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_m_2)

```


### Regression Based Dementia_3 

```{r}

summary(res_rb_m_3)

result_minority_crp_rb_3 = extract_decompo(res_rb_m_3)

ggcmest(res_rb_m_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_rb_m_3 <- cmsens(res_rb_m_3, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_m_3)

```


### Regression Based Dementia_4 

```{r}

summary(res_rb_m_4)

result_minority_crp_rb_4 = extract_decompo(res_rb_m_4)

ggcmest(res_rb_m_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_rb_m_4 <- cmsens(res_rb_m_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_m_4)

```

### Sensitivity G Method Dementia_1

```{r}

summary(res_gformula_m_l_1)

result_minority_crp_gformula_l_1 = extract_decompo(res_gformula_m_l_1)

ggcmest(res_gformula_m_l_1) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_1 <- cmsens(res_gformula_m_l_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_1)

```


### Sensitivity G Method Dementia_2

```{r}

summary(res_gformula_m_l_2)

result_minority_crp_gformula_l_2 = extract_decompo(res_gformula_m_l_2)

ggcmest(res_gformula_m_l_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_2 <- cmsens(res_gformula_m_l_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_2)

```


### Sensitivity G Method Dementia_3

```{r}

summary(res_gformula_m_l_3)

result_minority_crp_gformula_l_3 = extract_decompo(res_gformula_m_l_3)

ggcmest(res_gformula_m_l_3) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_3 <- cmsens(res_gformula_m_l_1, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_3)

```


### Sensitivity G Method Dementia_4

```{r}

summary(res_gformula_m_l_4)

result_minority_crp_gformula_l_4 = extract_decompo(res_gformula_m_l_4)

ggcmest(res_gformula_m_l_4) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_4 <- cmsens(res_gformula_m_l_4, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_4)

```


# Sensitivity G Method Education{.tabset}


## Minoritized Dementia

```{r}

# Minoritized 
library(MASS) # using the mass function you need to call directly the select fuction with dplyr, this library is needed to call the polr() function
library(Hmisc)

minority_data_dementia_edu <- dementia_incidence_6y %>% dplyr::select(dementia, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()

minority_data_dementia_edu <- minority_data_dementia_edu %>% 
                               mutate_at(vars(edu_cat), as.factor) 

minority_data_dementia_edu_2 <- dementia_incidence_6y %>% dplyr::select(dementia_2, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

minority_data_dementia_edu_2 <- minority_data_dementia_edu_2 %>% 
                              mutate_at(vars(edu_cat), as.factor) 

# Minority Model with multiple L (mediator-outcome) confounders affected by the exposure variable dementia 

res_gformula_m_l_edu <- cmest(data = minority_data_dementia_edu, model = "gformula", outcome = "dementia", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=minority_data_dementia_edu),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia_edu), polr(edu_cat ~ minority + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = minority_data_dementia_edu, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_m_l_edu)

result_minority_crp_gformula_l_edu = extract_decompo(res_gformula_m_l_edu)

ggcmest(res_gformula_m_l_edu) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_edu <- cmsens(res_gformula_m_l_edu, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_edu)


```


## Minoritized Dementia_2 


```{r}

# Minority Model with multiple L (mediator-outcome) confounders affected by the exposure variable dementia_2 


res_gformula_m_l_edu_2 <- cmest(data = minority_data_dementia_edu_2, model = "gformula", outcome = "dementia_2", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_2 ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=minority_data_dementia_edu_2),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = minority_data_dementia_edu_2), polr(edu_cat ~ minority + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = minority_data_dementia_edu_2, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_m_l_edu_2)

result_minority_crp_gformula_l_edu_2 = extract_decompo(res_gformula_m_l_edu_2)

ggcmest(res_gformula_m_l_edu_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l_edu_2 <- cmsens(res_gformula_m_l_edu_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l_edu_2)

```


## Black Dementia_1 

```{r}

# Black 

black_data_dementia_edu <- dementia_incidence_6y %>% dplyr::select(dementia, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()

black_data_dementia_edu <- black_data_dementia_edu %>% 
                               mutate_at(vars(edu_cat), as.factor) 

black_data_dementia_edu_2 <- dementia_incidence_6y %>% dplyr::select(dementia_2, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

black_data_dementia_edu_2 <- black_data_dementia_edu_2 %>% 
                               mutate_at(vars(edu_cat), as.factor) 


# black Model with multiple L (mediator-outcome) confounders affected by the exposure variable - Dementia 

res_gformula_black_l_edu <- cmest(data = black_data_dementia_edu, model = "gformula", outcome = "dementia", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ black + baseline_crp_75 + black*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=black_data_dementia_edu),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_edu), polr(edu_cat ~ black + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = black_data_dementia_edu, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_black_l_edu)

result_b_crp_gformula_l_edu = extract_decompo(res_gformula_black_l_edu)

ggcmest(res_gformula_black_l_edu) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_black_l_edu <- cmsens(res_gformula_black_l_edu, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_black_l_edu)


```



## Black Dementia_2 


```{r}


# black Model with multiple L (mediator-outcome) confounders affected by the exposure variable - Dementia_2 

res_gformula_black_l_edu_2 <- cmest(data = black_data_dementia_edu_2, model = "gformula", outcome = "dementia_2", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_2 ~ black + baseline_crp_75 + black*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=black_data_dementia_edu_2),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = black_data_dementia_edu_2), polr(edu_cat ~ black + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = black_data_dementia_edu_2, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_black_l_edu_2)

result_b_crp_gformula_l_edu_2 = extract_decompo(res_gformula_black_l_edu_2)

ggcmest(res_gformula_black_l_edu_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_black_l_edu_2 <- cmsens(res_gformula_black_l_edu_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_black_l_edu_2)


```


## Hispanic Dementia_1

```{r}
# Hispanic 

hispanic_data_dementia_edu <- dementia_incidence_6y %>% dplyr::select(dementia, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time) %>% na.omit()

hispanic_data_dementia_edu <- hispanic_data_dementia_edu %>% 
                               mutate_at(vars(edu_cat), as.factor) 

hispanic_data_dementia_edu_2 <- dementia_incidence_6y %>% dplyr::select(dementia_2, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat, z_bmi, z_alcohol, smoking_1, smoking_2, chronic, ln_time_2) %>% na.omit()

hispanic_data_dementia_edu_2 <- hispanic_data_dementia_edu_2 %>% 
                               mutate_at(vars(edu_cat), as.factor) 

# Hispanic 

# Hispanic Model with multiple L (mediator-outcome) confounders affected by the exposure variable - Dementia 
set.seed(2021)
res_gformula_hispanic_l_edu <- cmest(data = hispanic_data_dementia_edu, model = "gformula", outcome = "dementia", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time, family = poisson(link = "log"), data=hispanic_data_dementia_edu),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia_edu), polr(edu_cat ~ hispanic + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = hispanic_data_dementia_edu, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_hispanic_l_edu)

result_h_crp_gformula_l_edu = extract_decompo(res_gformula_hispanic_l_edu)

ggcmest(res_gformula_hispanic_l_edu) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_hispanic_l_edu <- cmsens(res_gformula_hispanic_l_edu, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_hispanic_l_edu)

```


## Hispanic Dementia_2 


```{r}

# Hispanic Model with multiple L (mediator-outcome) confounders affected by the exposure variable - Dementia_2 
set.seed(2021)
res_gformula_hispanic_l_edu_2 <- cmest(data = hispanic_data_dementia_edu_2, model = "gformula", outcome = "dementia_2", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat", "z_bmi", "z_alcohol", "smoking_1", "smoking_2", "chronic", "ln_time_2"), postc= c("APOE2010_bin", "edu_cat"), EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia_2 ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + factor(edu_cat) + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, offset = ln_time_2, family = poisson(link = "log"), data=hispanic_data_dementia_edu_2),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, family = binomial(link = "logit"), data = hispanic_data_dementia_edu_2), polr(edu_cat ~ hispanic + wave + z_age + sex+ z_bmi + z_alcohol + smoking_1 + smoking_2 + chronic, data = hispanic_data_dementia_edu_2, Hess=TRUE)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 3, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


summary(res_gformula_hispanic_l_edu_2)

result_h_crp_gformula_l_edu_2 = extract_decompo(res_gformula_hispanic_l_edu_2)

ggcmest(res_gformula_hispanic_l_edu_2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_hispanic_l_edu_2 <- cmsens(res_gformula_hispanic_l_edu_2, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_hispanic_l_edu_2)

```




# Table 3 & Supplemental Table 6


Mediation analysis results for tables 3 and supplemental table 6

```{r}

# save single dementia incident models blacks and hispanics 
single_mediation_dementia_black_crp_1 = rbind(result_black_crp_rb_1, result_black_crp_gformula_l_1)
single_mediation_dementia_black_crp_2 = rbind(result_black_crp_rb_2, result_black_crp_gformula_l_2)
single_mediation_dementia_black_crp_3 = rbind(result_black_crp_rb_3, result_black_crp_gformula_l_3)
single_mediation_dementia_hispanic_crp_1 = rbind(result_hispanic_crp_rb_1, result_hispanic_crp_gformula_l_1)
single_mediation_dementia_hispanic_crp_2 = rbind(result_hispanic_crp_rb_2, result_hispanic_crp_gformula_l_2)
single_mediation_dementia_hispanic_crp_3 = rbind(result_hispanic_crp_rb_3, result_hispanic_crp_gformula_l_3)
single_mediation_dementia_minority_crp_1 = rbind(result_minority_crp_rb_1, result_minority_crp_gformula_l_1)
single_mediation_dementia_minority_crp_2 = rbind(result_minority_crp_rb_2, result_minority_crp_gformula_l_2)
single_mediation_dementia_minority_crp_3 = rbind(result_minority_crp_rb_3, result_minority_crp_gformula_l_3)


sheets_IRR = list('single_med_dem_b_1' = single_mediation_dementia_black_crp_1, 'single_med_dem_h_1' = single_mediation_dementia_hispanic_crp_1, 'single_med_dem_m_1' = single_mediation_dementia_minority_crp_1)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'single_mediation_crp_1_9-6-23.xlsx'))

sheets_IRR = list('single_med_dem_b_2' = single_mediation_dementia_black_crp_2, 'single_med_dem_h_2' = single_mediation_dementia_hispanic_crp_2, 'single_med_dem_m_2' = single_mediation_dementia_minority_crp_2)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'single_mediation_crp_2_9-6-23.xlsx'))

sheets_IRR = list('single_med_dem_b_3' = single_mediation_dementia_black_crp_3, 'single_med_dem_h_3' = single_mediation_dementia_hispanic_crp_3, 'single_med_dem_m_3' = single_mediation_dementia_minority_crp_3)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'single_mediation_crp_3_9-6-23.xlsx'))


# save sensitivity models using education as L for each racialized social group

mediation_dementia_black_crp_edu = rbind(result_b_crp_gformula_l_edu, result_b_crp_gformula_l_edu_2)
mediation_dementia_hispanic_crp_edu = rbind(result_h_crp_gformula_l_edu, result_h_crp_gformula_l_edu_2)
mediation_dementia_minority_crp_edu = rbind(result_minority_crp_gformula_l_edu, result_minority_crp_gformula_l_edu_2)


sheets_IRR = list('med_dem_b_edu' = mediation_dementia_black_crp_edu, 'med_dem_h_edu' = mediation_dementia_hispanic_crp_edu, 'med_dem_m_edu' = mediation_dementia_minority_crp_edu)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'mediation_crp_edu_9-6-23.xlsx'))


```


# Supplemental Table 7 

```{r}

result_evalue_black_rb_1 = extract_evalue(evalue_rb_1)
result_evalue_hispanic_rb_1 = extract_evalue(evalue_rb_h_1)
result_evalue_minority_rb_1 = extract_evalue(evalue_rb_m_1)

result_evalue_black_rb_2 = extract_evalue(evalue_rb_2)
result_evalue_hispanic_rb_2 = extract_evalue(evalue_rb_h_2)
result_evalue_minority_rb_2 = extract_evalue(evalue_rb_m_2)

result_evalue_black_rb_3 = extract_evalue(evalue_rb_3)
result_evalue_hispanic_rb_3 = extract_evalue(evalue_rb_h_3)
result_evalue_minority_rb_3 = extract_evalue(evalue_rb_m_3)

result_evalue_black_gf_1 = extract_evalue(evalue_gformula_b_l_1)
result_evalue_hispanic_gf_1 = extract_evalue(evalue_gformula_h_l_1)
result_evalue_minority_gf_1  = extract_evalue(evalue_gformula_m_l_1)

result_evalue_black_gf_2 = extract_evalue(evalue_gformula_b_l_2)
result_evalue_hispanic_gf_2 = extract_evalue(evalue_gformula_h_l_2)
result_evalue_minority_gf_2  = extract_evalue(evalue_gformula_m_l_2)

result_evalue_black_gf_3 = extract_evalue(evalue_gformula_b_l_3)
result_evalue_hispanic_gf_3 = extract_evalue(evalue_gformula_h_l_3)
result_evalue_minority_gf_3  = extract_evalue(evalue_gformula_m_l_3)



evalue_black_1 = rbind(result_evalue_black_rb_1, result_evalue_black_gf_1)
evalue_black_2 = rbind(result_evalue_black_rb_2, result_evalue_black_gf_2)
evalue_black_3 = rbind(result_evalue_black_rb_3, result_evalue_black_gf_3)

evalue_hispanic_1 = rbind(result_evalue_hispanic_rb_1, result_evalue_hispanic_gf_1)
evalue_hispanic_2 = rbind(result_evalue_hispanic_rb_2, result_evalue_hispanic_gf_2)
evalue_hispanic_3 = rbind(result_evalue_hispanic_rb_3, result_evalue_hispanic_gf_3)


evalue_minority_1 = rbind(result_evalue_minority_rb_1, result_evalue_minority_gf_1)
evalue_minority_2 = rbind(result_evalue_minority_rb_2, result_evalue_minority_gf_2)
evalue_minority_3 = rbind(result_evalue_minority_rb_3, result_evalue_minority_gf_3)


sheets_evalue = list('evalue_b' = evalue_black_1, 'evalue_h' = evalue_hispanic_1, 'evalue_m' = evalue_minority_1)
write_xlsx(sheets_evalue, path = paste0(source_dir_data, 'evalue_1_crp_9-6-23.xlsx'))


sheets_evalue = list('evalue_b' = evalue_black_2, 'evalue_h' = evalue_hispanic_2, 'evalue_m' = evalue_minority_2)
write_xlsx(sheets_evalue, path = paste0(source_dir_data, 'evalue_2_crp_9-6-23.xlsx'))


sheets_evalue = list('evalue_b' = evalue_black_3, 'evalue_h' = evalue_hispanic_3, 'evalue_m' = evalue_minority_3)
write_xlsx(sheets_evalue, path = paste0(source_dir_data, 'evalue_3_crp_9-6-23.xlsx'))


```



# Suplemental Table 11

```{r}

# save mediation analysis 
mediation_dementia_black_crp_log = rbind(result_black_crp_rb_4, result_black_crp_gformula_l_4)
mediation_dementia_hispanic_crp_log = rbind(result_hispanic_crp_rb_4, result_hispanic_crp_gformula_l_4)
mediation_dementia_minority_crp_log = rbind(result_minority_crp_rb_4, result_minority_crp_gformula_l_4)



sheets_IRR = list('med_dem_b_log_crp' = mediation_dementia_black_crp_log, 'med_dem_h_log_crp' = mediation_dementia_hispanic_crp_log, 'med_dem_m_log_crp' = mediation_dementia_minority_crp_log)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'mediation_log_crp_1-13-24.xlsx'))

```



# Supplemental Figure 3 

Supplemental Figure 3: Plot of mediation analysis estimates from regression-based and randomized analogue models denoting percent of the racial disparity in incident dementia that is due to the mediating effect of C-reactive protein, the percent due to the interaction between exposure and mediator, and the proportion eliminated. A. Mediation estimates from the racial disparity between the minoritized racial group (non-Hispanic Black and Hispanic) vs the non-Hispanic White group. B. Mediation estimates from the racial disparity between the non-Hispanic Black vs the non-Hispanic White group. C. Mediation estimates from the racial disparity between the Hispanic group vs the non-Hispanic White group.



```{r}

library(ggplot2)
library(ggpubr)
theme_set(theme_pubr()) 

library(readxl)
estimates <- read_excel("/Users/cesarhiggins/Documents/CRP_Dementia_8-9-23/Mediation Results/estimates_9-6-23.xlsx", "Black")
  

estimates_h <- read_excel("/Users/cesarhiggins/Documents/CRP_Dementia_8-9-23/Mediation Results/estimates_9-6-23.xlsx", "Hispanic")


estimates_m <- read_excel("/Users/cesarhiggins/Documents/CRP_Dementia_8-9-23/Mediation Results/estimates_9-6-23.xlsx", "Minoritized")


p_black <- ggplot(data=estimates,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - non-Hispanic Black vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
   coord_flip() 

p_black

 
p_hispanic <- ggplot(data=estimates_h,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - Hispanic vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    coord_flip()

p_hispanic

 
p_minoritized <- ggplot(data=estimates_m,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - Minoritized vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    coord_flip()

p_minoritized




 p_b_h <- ggarrange(
  p_minoritized, p_black, p_hispanic, labels = c("A", "B", "C"),
  common.legend = TRUE, legend = "bottom"
  )
 
 p_b_h
```