---
title: "Mediation CRP Analysis"
author: "César Higgins"
date: '2023-01-19'
output:
  html_document: null
  highlight: pygments
  word_document: default
  keep_md: no
  number_sections: yes
  smart: yes
  theme: sandstone
  toc: yes
  toc_float: yes
  code_folding: hide
---


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(compareGroups)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(lubridate)
library(magrittr) # need to run every time you start R and want to use %>%
library(kableExtra)
library(gtsummary)
library(foreign)
library(gmodels)
library(stringr)
library(dplyr)  
library(haven)

summarize = dplyr::summarize

knitr::opts_chunk$set(
  eval       = TRUE,    # whether to run code in code chunk
  include    = TRUE,    # whether to include the chunk output
  echo       = TRUE,   # Whether to show code chunk in final output
  error      = TRUE,    # whether to display error messages
  message    = FALSE,   # whether to preserve messages
  warning    = FALSE,   # whether to preserve warnings
  comment    = "#>",    # a character string to append at start
                        # of each line of results in final document
  tidy       = FALSE,   # whether to tidy code chunks for display
  #dpi        = 96, 
  #fig.width  = 6,       # consistent width for figures
  #fig.asp    = 0.618,   # the golden ratio, can be adjusted in individual chunks
  out.width  = "100%",   # controls the output size
  fig.align  = "center" # give plot room to breathe
)

```




# Mediation 

```{r}

library(haven)
# import data from stata for mediation analysis
rm(list = ls())
source_dir_dta <- "/Users/cesarhiggins/Documents/Race_CRP_Dementia/STATA/"
dementia_incidence_6y <- read_dta(paste0(source_dir_dta, "crp_data_incidence_mediation.dta")) # file with analytic sample size


library(CMAverse)

```


```{r, results='hide'}

library(writexl)
source_dir_data <- "/Users/cesarhiggins/Documents/Race_CRP_Dementia/Mediation Results/"

# function to extract decomposition components 
# Link to existing files for function use
source(paste0(source_dir_data, "extract_decompo.R"))
source(paste0(source_dir_data, "extract_evalue.R"))

```



# Black Models

```{r, results='hide'}

black_data <- dementia_incidence_6y %>% select(dementia, black, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoke, chronic, ln_time) %>% na.omit()



# Regression Based Approach
set.seed(2021)
res_rb_b <- cmest(data = black_data, model = "rb", outcome = "dementia", exposure = "black",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), EMint=TRUE,
                mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, family = binomial(link = "logit"), data = black_data)), yreg = glm(dementia ~ black + baseline_crp_75 + black*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, offset = ln_time, family = poisson(link = "log"), data=black_data),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)



# G estimation L approach 

set.seed(2021)
res_gformula_b_l <- cmest(data = black_data, model = "gformula", outcome = "dementia", exposure = "black",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list(glm(baseline_crp_75 ~ black + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, family = binomial(link = "logit"), data = black_data)), yreg = glm(dementia ~ black + baseline_crp_75 + black*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, , offset = ln_time, family = poisson(link = "log"), data=black_data),  postcreg = list(glm(APOE2010_bin ~ black + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, family = binomial(link = "logit"), data = black_data)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)




```

## hsCRP Mediation Incident Dementia Blacks Models{.tabset}


### Regression Based 

```{r}

summary(res_rb_b)

result_black_crp_rb = extract_decompo(res_rb_b)

ggcmest(res_rb_b) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_rb <- cmsens(res_rb_b, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb)

```


### Sensitivity G Method 

```{r}

summary(res_gformula_b_l)

result_black_crp_gformula_l = extract_decompo(res_gformula_b_l)

ggcmest(res_gformula_b_l) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_gformula_b_l <- cmsens(res_gformula_b_l, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_b_l)


```


# Hispanic Models


```{r, results='hide'}

hispanic_data <- dementia_incidence_6y %>% select(dementia, hispanic, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoke, chronic, ln_time) %>% na.omit()

set.seed(2021)

# Regression Based Approach

res_rb_h <- cmest(data = hispanic_data, model = "rb", outcome = "dementia", exposure = "hispanic",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, , offset = ln_time, family = poisson(link = "log"), data=hispanic_data),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)


# G estimation L approach 

set.seed(2021)
res_gformula_h_l <- cmest(data = hispanic_data, model = "gformula", outcome = "dementia", exposure = "hispanic",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ hispanic + baseline_crp_75 + hispanic*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, , offset = ln_time, family = poisson(link = "log"), data=hispanic_data),  postcreg = list(glm(APOE2010_bin ~ hispanic + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, family = binomial(link = "logit"), data = hispanic_data)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)


```


## hsCRP Mediation Incident Dementia Hispanic Models{.tabset}



### Regression Based 


```{r}

summary(res_rb_h)

result_hispanic_crp_rb = extract_decompo(res_rb_h)

ggcmest(res_rb_h) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))



evalue_rb_h <- cmsens(res_rb_h, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_h)

```



### Sensitivity G Method 

```{r}

summary(res_gformula_h_l)

result_hispanic_crp_gformula_l = extract_decompo(res_gformula_h_l)

ggcmest(res_gformula_h_l) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_h_l <- cmsens(res_gformula_h_l, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_h_l)

```


# Minoritized Models 


```{r, results='hide'}

minority_data <- dementia_incidence_6y %>% select(dementia, minority, baseline_crp_75, wave, APOE2010_bin, z_age, sex, edu_cat_1, edu_cat_2, z_bmi, z_alcohol, smoke, chronic, ln_time) %>% na.omit()



# Regression Based Approach
set.seed(2021)
res_rb_m <- cmest(data = minority_data, model = "rb", outcome = "dementia", exposure = "minority",
                mediator = c("baseline_crp_75"), basec = c("wave", "APOE2010_bin", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), EMint = TRUE,
                mreg = list("logistic"), yreg = glm(dementia ~ minority + baseline_crp_75 + minority*baseline_crp_75+ wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, , offset = ln_time, family = poisson(link = "log"), data=minority_data),
                astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 0, 2, 0, 0, 0, 0, 0, 0), yval = list(0), estimation = "imputation", inference = "bootstrap", nboot = 1000, multimp = FALSE)



# G estimation L approach 

set.seed(2021)
res_gformula_m_l <- cmest(data = minority_data, model = "gformula", outcome = "dementia", exposure = "minority",
                      mediator = c("baseline_crp_75"), basec = c("wave", "z_age", "sex", "edu_cat_1", "edu_cat_2" , "z_bmi", "z_alcohol", "smoke", "chronic", "ln_time"), postc= "APOE2010_bin", EMint = TRUE,
                      mreg = list("logistic"), yreg = glm(dementia ~ minority + baseline_crp_75 + minority*baseline_crp_75 + wave + APOE2010_bin + z_age + sex + edu_cat_1 + edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, , offset = ln_time, family = poisson(link = "log"), data=minority_data),  postcreg = list(glm(APOE2010_bin ~ minority + wave + z_age + sex + edu_cat_1 + 
    edu_cat_2 + z_bmi + z_alcohol + smoke + chronic, family = binomial(link = "logit"), data = minority_data)),
                      astar = 0, a = 1, mval = list(0), basecval = list(2006, 0, 2, 0, 0, 0, 0, 0, 0), yval=list(0),
                     estimation = "imputation", inference = "bootstrap", nboot = 1000)

```


## hsCRP Mediation Incident Dementia Minority Models{.tabset}

### Regression Based 

```{r}

summary(res_rb_m)

result_minority_crp_rb = extract_decompo(res_rb_m)

ggcmest(res_rb_m) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))


evalue_rb_m <- cmsens(res_rb_m, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_rb_m)

```


### Sensitivity G Method 

```{r}

summary(res_gformula_m_l)

result_minority_crp_gformula_l = extract_decompo(res_gformula_m_l)

ggcmest(res_gformula_m_l) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))

evalue_gformula_m_l <- cmsens(res_gformula_m_l, sens = "uc",   MEmethod = "simex", MEvariable = NULL, MEvartype = NULL, MEerror = NULL, lambda = c(0.5, 1, 1.5, 2), B = 200, nboot.rc = 1000)

print(evalue_gformula_m_l)

```


# Table 3 & Supplemental Table 6


Mediation analysis results for tables 3 and supplemental table 6

```{r}

# save single dementia incident models blacks and hispanics 
single_mediation_dementia_black_crp = rbind(result_black_crp_rb, result_black_crp_gformula_l)
single_mediation_dementia_hispanic_crp = rbind(result_hispanic_crp_rb, result_hispanic_crp_gformula_l)
single_mediation_dementia_minority_crp = rbind(result_minority_crp_rb, result_minority_crp_gformula_l)
sheets_IRR = list('single_med_dem_b' = single_mediation_dementia_black_crp, 'single_med_dem_h' = single_mediation_dementia_hispanic_crp, 'single_med_dem_m' = single_mediation_dementia_minority_crp)
write_xlsx(sheets_IRR, path = paste0(source_dir_data, 'single_mediation_crp_2-10-23.xlsx'))



```


# Supplemental Table 7 

```{r}

result_evalue_black_rb = extract_evalue(evalue_rb)
result_evalue_hispanic_rb = extract_evalue(evalue_rb_h)
result_evalue_minority_rb = extract_evalue(evalue_rb_m)

result_evalue_black_gf = extract_evalue(evalue_gformula_b_l)
result_evalue_hispanic_gf = extract_evalue(evalue_gformula_h_l)
result_evalue_minority_gf = extract_evalue(evalue_gformula_m_l)


evalue_black = rbind(result_evalue_black_rb, result_evalue_black_gf)
evalue_hispanic = rbind(result_evalue_hispanic_rb, result_evalue_hispanic_gf)
evalue_minority = rbind(result_evalue_minority_rb, result_evalue_minority_gf)

sheets_evalue = list('evalue_b' = evalue_black, 'evalue_h' = evalue_hispanic, 'evalue_m' = evalue_minority)
write_xlsx(sheets_evalue, path = paste0(source_dir_data, 'evalue_crp_2-10-23.xlsx'))

```



# Supplemental Figure 3 

Supplemental Figure 3: Plot of mediation analysis estimates from regression-based and randomized analogue models denoting percent of the racial disparity in incident dementia that is due to the mediating effect of C-reactive protein, the percent due to the interaction between exposure and mediator, and the proportion eliminated. A. Mediation estimates from the racial disparity between the minoritized racial group (non-Hispanic Black and Hispanic) vs the non-Hispanic White group. B. Mediation estimates from the racial disparity between the non-Hispanic Black vs the non-Hispanic White group. C. Mediation estimates from the racial disparity between the Hispanic group vs the non-Hispanic White group.



```{r}

library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

library(readxl)
estimates <- read_excel("/Users/cesarhiggins/Documents/CRP Dementia/Mediation Results/estimates.xlsx", "Black")
  

estimates_h <- read_excel("/Users/cesarhiggins/Documents/CRP Dementia/Mediation Results/estimates.xlsx", "Hispanic")


estimates_m <- read_excel("/Users/cesarhiggins/Documents/CRP Dementia/Mediation Results/estimates.xlsx", "Minoritized")


p_black <- ggplot(data=estimates,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - non-Hispanic Black vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
   coord_flip() 

p_black

 
p_hispanic <- ggplot(data=estimates_h,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - Hispanic vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    coord_flip()

p_hispanic

 
p_minoritized <- ggplot(data=estimates_m,
    aes(x = Effect,y = Estimate, ymin = Lower, ymax = Upper))+
    geom_pointrange(aes(col=Effect))+
    geom_hline(aes(fill=Effect),yintercept =c(0,1), linetype=2)+
    ggtitle("Estimates (%) - Minoritized vs non-Hispanic White") +
    xlab('Models')+ ylab("Effect Estimates (95% Confidence Interval)")+
    geom_errorbar(aes(ymin=Lower, ymax=Upper,col=Effect),width=0.5,cex=1)+ 
    facet_wrap(~Model,strip.position="left",nrow=6,scales = "free_y") +
    theme(plot.title=element_text(size=16,face="bold"),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_text(size=12,face="bold"),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"))+
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    coord_flip()

p_minoritized




 p_b_h <- ggarrange(
  p_minoritized, p_black, p_hispanic, labels = c("A", "B", "C"),
  common.legend = TRUE, legend = "bottom"
  )
 
 p_b_h
```